---
title: "Simulation Results: Fixed Sample Size"
subtitle: "Moose Abundance Estimation"
date: "`r Sys.Date()`"
author: "Péter Sólymos"
toc: true
format:
  html: 
    html-math-method: katex
    self-contained: true
  pdf: default
---

TODO:

- remove ZINB, ZIP
- add CI to the coverage table (Gassaway too)
- figure out NB issue (coverage is too low, try different theta)
- BIC pairwise plots

Gassaway:

- SRL to check variance formula
- RSF to use for "true" classification, take top 30% high
- Do binomial: 10% flips (independent of RSF)
- Check N1/N2 for Gassaway
- Drop the Normal version

Simulated cells using Sophie's South Canol RSF data, 
sampled 100 of the cells randomly (same as in the real survey). 
This was repeated 200 times for each of the 6 models (P, NB, HP, HNB, ZIP, ZINB). 
Used 500 bootstrap iterations.

The same 6 models were fit to the data, then the following metrics were saved:

- `true_total`: sum of counts for all cells
- total abundance (`mean`) estimate: cell level estimates summed for each bootstrap run, boostrap mean calculated
- 90% and 95% PIs for total abundance (quantiles of the bootstrap total counts)
- coverage (90% and 95%): proportion of unsurveyed cells where the true count was inside the cell level PIs

We also saved coefficients (true and estimated) and model selection results (AIC, BIC).

```{r}
#| include: false
library(knitr)
library(intrval)
library(ggplot2)
library(dplyr)

devtools::load_all("../..")
DIR <- "../../_tmp/simuls"
# devtools::load_all()
# DIR <- "_tmp/simuls"

fl <- list.files(DIR)
dd0 <- NULL
for (fi in fl) {
    MODEL <- gsub("scanol-", "", strsplit(fi, "_")[[1]][2])
    fn <- paste0(DIR, "/", fi)
    load(fn)

    d <- do.call(
        rbind,
        lapply(RES, \(x) {
            z <- data.frame(
                true_model = MODEL,
                model = colnames(x),
                as.data.frame(t(x))
            )
        })
    )
    d$pi_95 <- as.numeric(d[, "true_total"] %[]% d[, c("X2.5.", "X97.5.")])
    d$pi_90 <- as.numeric(d[, "true_total"] %[]% d[, c("X5.", "X95.")])
    d$bias <- d$mean - d$true_total
    d$rel_bias <- (d$mean - d$true_total) / d$true_total
    colnames(d)[colnames(d) == "X2.5."] <- "ci_95_lower"
    colnames(d)[colnames(d) == "X5."] <- "ci_90_lower"
    colnames(d)[colnames(d) == "X95."] <- "ci_90_upper"
    colnames(d)[colnames(d) == "X97.5."] <- "ci_95_upper"

    dd0 <- rbind(dd0, d)
}
dd0$misclass <- NA_character_

ci_fun <- function(tau_hat, var_hat, level = 0.95) {
    a <- (1 - level) / 2
    a <- c(a, 1 - a)
    tau_hat + sqrt(var_hat) * qnorm(a)
}

GR <- NULL
for (fi in fl) {
    MODEL <- gsub("scanol-", "", strsplit(fi, "_")[[1]][2])
    fn <- paste0(DIR, "/", fi)
    load(fn)

    for (i in 1:ncol(MT)) {
        v <- quantile(m0$z, 0.75)
        m <- m0 |> sim_strat(threshold = v, p_misclass = 0.1)
        m$stratum0 <- ifelse(m$z >= v, 1, 0)
        summary(m)
        table(true = m$stratum0, miscl = m$stratum)

        g0 <- mc_gassaway(
            y1 = MT[SRV[, i] == 1 & m$stratum0 == 0, i],
            y2 = MT[SRV[, i] == 1 & m$stratum0 == 1, i],
            N1 = sum(1 - m$stratum0),
            N2 = sum(m$stratum0)
        )
        g10 <- mc_gassaway(
            y1 = MT[SRV[, i] == 1 & m$stratum == 0, i],
            y2 = MT[SRV[, i] == 1 & m$stratum == 1, i],
            N1 = sum(1 - m$stratum),
            N2 = sum(m$stratum)
        )
        tot <- sum(MT[, i])
        ci0_90 <- ci_fun(g0$total, g0$variance, 0.90)
        ci0_95 <- ci_fun(g0$total, g0$variance, 0.95)
        gr0 <- data.frame(
            true_model = MODEL,
            misclass = "none",
            model = "G",
            true_total = tot,
            mean = g0$total,
            ci_90_lower = ci0_90[1],
            ci_90_upper = ci0_90[2],
            ci_95_lower = ci0_95[1],
            ci_95_upper = ci0_95[2],
            coverage_90 = ifelse(tot %[]% ci0_90, 1, 0),
            coverage_95 = ifelse(tot %[]% ci0_95, 1, 0)
        )
        ci10_90 <- ci_fun(g10$total, g10$variance, 0.90)
        ci10_95 <- ci_fun(g10$total, g10$variance, 0.95)
        gr10 <- data.frame(
            true_model = MODEL,
            misclass = "B(0.1)",
            model = "G",
            true_total = tot,
            mean = g10$total,
            ci_90_lower = ci10_90[1],
            ci_90_upper = ci10_90[2],
            ci_95_lower = ci10_95[1],
            ci_95_upper = ci10_95[2],
            coverage_90 = ifelse(tot %[]% ci10_90, 1, 0),
            coverage_95 = ifelse(tot %[]% ci10_95, 1, 0)
        )
        GR <- rbind(GR, rbind(gr0, gr10))
    }
}
GR$bias <- GR$mean - GR$true_total
GR$rel_bias <- (GR$mean - GR$true_total) / GR$true_total
GR$median <- NA_real_
GR$mode <- NA_real_
GR$pi_95 <- NA_real_
GR$pi_90 <- NA_real_

dd <- rbind(dd0, GR)
# summary(dd)

MS <- NULL
cn <- "BIC"
rn <- c("HNB", "HP", "NB", "P", "ZINB", "ZIP")
for (fi in fl) {
    MODEL <- gsub("scanol-", "", strsplit(fi, "_")[[1]][2])
    fn <- paste0(DIR, "/", fi)
    load(fn)

    ms <- t(sapply(IC, \(z) {
        z[rn, cn]
    }))
    colnames(ms) <- rn
    sel <- character()
    rnk <- numeric()
    delta <- numeric()
    for (k in 1:nrow(ms)) {
        r <- rank(ms[k, ])
        sel[k] <- names(r)[r == 1]
        rnk[k] <- unname(r[MODEL])
        delta[k] <- ms[k, MODEL] - ms[k, sel[k]]
    }
    ms <- data.frame(true_model = MODEL, best_model = sel, true_model_rank = rnk, delta = delta, ms)

    MS <- rbind(MS, ms)
}
# summary(MS)
```

The true total abundance varied somewhat between the setups, might have to tune a bit more to make them more comparable:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = true_total)) +
    geom_boxplot() +
    theme_light()
```


The total moose abundance coverage was lower than nominal for all models
(95% confidence intervals presented in parentheses):

```{r}
#| echo: false
#| output: asis
df <- dd |>
    filter(model != "G") |>
    group_by(true_model, model) |>
    summarize(pi_90 = mean(pi_90), pi_95 = mean(pi_95), .groups = "keep") |>
    mutate(
        phat_pm_90 = 1.96 * sqrt(pi_90 * (1 - pi_90) / 200),
        phat_pm_95 = 1.96 * sqrt(pi_95 * (1 - pi_95) / 200),
        ci_lo_90 = round(pi_90 - phat_pm_90, 3),
        ci_hi_90 = round(pi_90 + phat_pm_90, 3),
        ci_lo_95 = round(pi_95 - phat_pm_95, 3),
        ci_hi_95 = round(pi_95 + phat_pm_95, 3),
        pi_90 = paste0(pi_90, " (", ci_lo_90, "--", ci_hi_90, ")"),
        pi_95 = paste0(pi_95, " (", ci_lo_95, "--", ci_hi_95, ")")
    ) |>
    select(-phat_pm_90, -phat_pm_95, -ci_lo_90, -ci_hi_90, -ci_lo_95, -ci_hi_95)
knitr::kable(df)
```

The bias (estimate - true) in total abundance:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = bias, col = model)) +
    geom_hline(yintercept = 0) +
    geom_boxplot() +
    theme_light()
```

Relative bias (standardized by true abundance):

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = rel_bias, col = model)) +
    geom_hline(yintercept = 0) +
    geom_boxplot() +
    theme_light()
```

Cell level coverage was higher than nominal:

```{r}
#| echo: false
dd |>
    filter(model != "G") |>
    ggplot(aes(x = true_model, y = coverage_95, col = model)) +
    geom_hline(yintercept = 0.95) +
    geom_boxplot() +
    theme_light()
dd |>
    filter(model != "G") |>
    ggplot(aes(x = true_model, y = coverage_90, col = model)) +
    geom_hline(yintercept = 0.90) +
    geom_boxplot() +
    theme_light()
```

Gassaway

```{r}
#| echo: false
dd |>
    filter(model == "G") |>
    ggplot(aes(x = true_model, y = bias, col = misclass)) +
    geom_hline(yintercept = 0) +
    geom_boxplot() +
    theme_light()

dd |>
    filter(model == "G") |>
    ggplot(aes(x = true_model, y = rel_bias, col = misclass)) +
    geom_hline(yintercept = 0) +
    geom_boxplot() +
    theme_light()
```

```{r}
#| echo: false
#| output: asis
df <- dd |>
    filter(model == "G") |>
    group_by(true_model, misclass) |>
    summarize(pi_90 = mean(coverage_90), pi_95 = mean(coverage_95), .groups = "keep") |>
    mutate(
        phat_pm_90 = 1.96 * sqrt(pi_90 * (1 - pi_90) / 200),
        phat_pm_95 = 1.96 * sqrt(pi_95 * (1 - pi_95) / 200),
        ci_lo_90 = round(pi_90 - phat_pm_90, 3),
        ci_hi_90 = round(pi_90 + phat_pm_90, 3),
        ci_lo_95 = round(pi_95 - phat_pm_95, 3),
        ci_hi_95 = round(pi_95 + phat_pm_95, 3),
        pi_90 = paste0(pi_90, " (", ci_lo_90, "--", ci_hi_90, ")"),
        pi_95 = paste0(pi_95, " (", ci_lo_95, "--", ci_hi_95, ")")
    ) |>
    select(-phat_pm_90, -phat_pm_95, -ci_lo_90, -ci_hi_90, -ci_lo_95, -ci_hi_95) |>
    arrange(true_model, desc(misclass))
knitr::kable(df)
```

Model selection results: frequency of selecting the true model based on BIC

```{r}
#| echo: false
table(true_model = MS$true_model, best_model = MS$best_model)
```

The rank of the true model based on BIC

```{r}
#| echo: false
table(MS$true_model, MS$true_model_rank)
```

NOTE: pivot longer given true model X, show faceted plots

```{r}
#| echo: false
cn <- c("HNB", "HP", "NB", "P", "ZINB", "ZIP")
mmm <- NULL
for (i in 1:nrow(MS)) {
    tmp <- unlist(MS[i, cn])
    mm <- data.frame(true_model = MS$true_model[i], model = cn, BIC_true = unname(tmp[MS$true_model[i]]), BIC = unname(tmp))
    mmm <- rbind(mmm, mm)
}

mmm |>
    filter(true_model == "P") |>
    ggplot(aes(x = BIC_true, y = BIC)) +
    geom_abline(col = 4, lty = 2) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    theme_light() +
    xlab("BIC [true model]") +
    ylab("BIC [fitted model]") +
    labs(title = "True model: P") +
    facet_wrap(vars(model), scales = "free")

mmm |>
    filter(true_model == "NB") |>
    ggplot(aes(x = BIC_true, y = BIC)) +
    geom_abline(col = 4, lty = 2) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    theme_light() +
    xlab("BIC [true model]") +
    ylab("BIC [fitted model]") +
    labs(title = "True model: NB") +
    facet_wrap(vars(model), scales = "free")

mmm |>
    filter(true_model == "HP") |>
    ggplot(aes(x = BIC_true, y = BIC)) +
    geom_abline(col = 4, lty = 2) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    theme_light() +
    xlab("BIC [true model]") +
    ylab("BIC [fitted model]") +
    labs(title = "True model: HP") +
    facet_wrap(vars(model), scales = "free")

mmm |>
    filter(true_model == "HNB") |>
    ggplot(aes(x = BIC_true, y = BIC)) +
    geom_abline(col = 4, lty = 2) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    theme_light() +
    xlab("BIC [true model]") +
    ylab("BIC [fitted model]") +
    labs(title = "True model: HNB") +
    facet_wrap(vars(model), scales = "free")

mmm |>
    filter(true_model == "ZIP") |>
    ggplot(aes(x = BIC_true, y = BIC)) +
    geom_abline(col = 4, lty = 2) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    theme_light() +
    xlab("BIC [true model]") +
    ylab("BIC [fitted model]") +
    labs(title = "True model: ZIP") +
    facet_wrap(vars(model), scales = "free")

mmm |>
    filter(true_model == "ZINB") |>
    ggplot(aes(x = BIC_true, y = BIC)) +
    geom_abline(col = 4, lty = 2) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
    theme_light() +
    xlab("BIC [true model]") +
    ylab("BIC [fitted model]") +
    labs(title = "True model: ZINB") +
    facet_wrap(vars(model), scales = "free")

```

