---
title: "Simulation Results: Fixed Sample Size"
subtitle: "Moose Abundance Estimation"
date: "`r Sys.Date()`"
author: "Péter Sólymos"
toc: true
format:
  html: 
    html-math-method: katex
    self-contained: true
  pdf: default
---

Simulated 2500 cells, sampled 500 of them randomly. This was repeated 100
times for each of the 4 models (P, NB, ZIP, ZINB). Used 500 bootstrap iterations.

The same 4 models were fit to the data, then the following metrics were saved:

- `true_total`: sum of counts for all 2500 cells
- total abundance (`mean`) estimate: cell level estimates summed for each bootstrap run, boostrap mean calculated
- 90% and 95% PIs for total abundance (quantiles of the bootstrap total counts)
- coverage (90% and 95%): proportion of unsurveyed cells where the true count was inside the cell level PIs

```{r}
#| include: false
library(knitr)
library(intrval)
library(ggplot2)
library(dplyr)
devtools::load_all("../..")

DIR <- "../../_tmp/simuls"
dd <- NULL
for (MODEL in c("P", "NB", "ZIP", "ZINB")) {
    fn <- sprintf("%s/moose-sim_model-%s_n-500_B-500.RData", DIR, MODEL)
    load(fn)

    d <- do.call(
        rbind,
        lapply(RES, \(x) {
            z <- data.frame(
                true_model = MODEL,
                model = colnames(x),
                as.data.frame(t(x))
            )
        })
    )
    d$pi_95 <- as.numeric(d[, "true_total"] %[]% d[, c("X2.5.", "X97.5.")])
    d$pi_90 <- as.numeric(d[, "true_total"] %[]% d[, c("X5.", "X95.")])
    d$bias <- d$mean - d$true_total
    d$rel_bias <- (d$mean - d$true_total) / d$true_total

    dd <- rbind(dd, d)
}
```

The true total abundance varied somewhat between the setups, might have to tune a bit more to make them more comparable:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = true_total, col = model)) +
    geom_boxplot() +
    theme_light()
```


The total moose abundance coverage was lower than nominal for all models:

```{r}
#| echo: false
#| output: asis
dd |>
    group_by(true_model, model) |>
    summarize(pi_90 = mean(pi_90), pi_95 = mean(pi_95), .groups = "keep") |>
    knitr::kable()
```

The bias (estimate - true) in total abundance:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = bias, col = model)) +
    geom_boxplot() +
    theme_light()
```

Relative bias (standardized by true abundance):

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = rel_bias, col = model)) +
    geom_boxplot() +
    theme_light()
```

Cell level coverage was higher than nominal:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = coverage_95, col = model)) +
    geom_boxplot() +
    theme_light()
dd |>
    ggplot(aes(x = true_model, y = coverage_90, col = model)) +
    geom_boxplot() +
    theme_light()
```

Gassaway

```{r}
#| echo: false

ci_fun <- function(tau_hat, var_hat, level = 0.95) {
    a <- (1 - level) / 2
    a <- c(a, 1 - a)
    tau_hat + sqrt(var_hat) * qnorm(a)
}

GR <- NULL
for (MODEL in c("P", "NB", "ZIP", "ZINB")) {
    fn <- sprintf("%s/moose-sim_model-%s_n-500_B-500.RData", DIR, MODEL)
    load(fn)

    for (i in 1:ncol(MT)) {
        v <- 0
        m <- m0 |> sim_strat(threshold = v, p_misclass = 0.1)
        m$stratum0 <- ifelse(m$z >= v, 1, 0)
        g0 <- mc_gassaway(
            y1 = MT[SRV[, i] == 1 & m$stratum0 == 0, i],
            y2 = MT[SRV[, i] == 1 & m$stratum0 == 1, i],
            N1 = sum(1 - m$stratum0),
            N2 = sum(m$stratum0)
        )
        g10 <- mc_gassaway(
            y1 = MT[SRV[, i] == 1 & m$stratum == 0, i],
            y2 = MT[SRV[, i] == 1 & m$stratum == 1, i],
            N1 = sum(1 - m$stratum),
            N2 = sum(m$stratum)
        )
        tot <- sum(MT[, i])
        ci0_90 <- ci_fun(g0$total, g0$variance, 0.90)
        ci0_95 <- ci_fun(g0$total, g0$variance, 0.95)
        ci10_90 <- ci_fun(g10$total, g10$variance, 0.90)
        ci10_95 <- ci_fun(g10$total, g10$variance, 0.95)
        gr0 <- data.frame(
            true_model = MODEL,
            misclass = 0,
            true_total = tot,
            total = g0$total,
            ci_90_lower = ci0_90[1], ci_90_upper = ci0_90[2],
            ci_95_lower = ci0_95[1], ci_95_upper = ci0_95[2],
            coverage_90 = ifelse(tot %[]% ci0_90, 1, 0),
            coverage_95 = ifelse(tot %[]% ci0_90, 1, 0)
        )
        gr10 <- data.frame(
            true_model = MODEL,
            misclass = 0.1,
            true_total = tot,
            total = g10$total,
            ci_90_lower = ci10_90[1], ci_90_upper = ci10_90[2],
            ci_95_lower = ci10_95[1], ci_95_upper = ci10_95[2],
            coverage_90 = ifelse(tot %[]% ci10_90, 1, 0),
            coverage_95 = ifelse(tot %[]% ci10_90, 1, 0)
        )
        GR <- rbind(GR, rbind(gr0, gr10))
    }
}
GR$bias <- GR$total - GR$true_total
GR$rel_bias <- (GR$total - GR$true_total) / GR$true_total

```

```{r}
#| echo: false
GR |>
    ggplot(aes(x = true_model, y = bias, col = paste(100 * misclass, "%"))) +
    geom_boxplot() +
    theme_light()

GR |>
    ggplot(aes(x = true_model, y = rel_bias, col = paste(100 * misclass, "%"))) +
    geom_boxplot() +
    theme_light()
```

```{r}
#| echo: false
#| output: asis
GR |>
    group_by(true_model, misclass) |>
    summarize(pi_90 = mean(coverage_90), pi_95 = mean(coverage_95), .groups = "keep") |>
    knitr::kable()
```
