---
title: "Simulation Results: Fixed Sample Size"
subtitle: "Moose Abundance Estimation"
date: "`r Sys.Date()`"
author: "Péter Sólymos"
toc: true
format:
  html: 
    html-math-method: katex
    self-contained: true
  pdf: default
---

Simulated 250 cells, sampled 250 of them randomly. This was repeated 100
times for each of the 4 models (P, NB, ZIP, ZINB). Used 500 bootstrap iterations.

The same 4 models were fit to the data, then the following metrics were saved:

- `true_total`: sum of counts for all 2500 cells
- total abundance (`mean`) estimate: cell level estimates summed for each bootstrap run, boostrap mean calculated
- 90% and 95% PIs for total abundance (quantiles of the bootstrap total counts)
- coverage (90% and 95%): proportion of unsurveyed cells where the true count was inside the cell level PIs

```{r}
#| include: false
library(knitr)
library(intrval)
library(ggplot2)
library(dplyr)

DIR <- "../../_tmp/simuls"
dd <- NULL
for (MODEL in c("P", "NB", "ZIP", "ZINB")) {
    fn <- sprintf("%s/moose-sim_model-%s_B-500.RData", DIR, MODEL)
    load(fn)

    d <- do.call(
        rbind,
        lapply(RES, \(x) {
            z <- data.frame(
                true_model = MODEL,
                model = colnames(x),
                as.data.frame(t(x))
            )
        })
    )
    d$pi_95 <- as.numeric(d[, "true_total"] %[]% d[, c("X2.5.", "X97.5.")])
    d$pi_90 <- as.numeric(d[, "true_total"] %[]% d[, c("X5.", "X95.")])
    d$bias <- d$mean - d$true_total
    d$rel_bias <- (d$mean - d$true_total) / d$true_total

    dd <- rbind(dd, d)
}
```

The true total abundance varied somewhat between the setups, might have to tune a bit more to make them more comparable:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = true_total, col = model)) +
    geom_boxplot() +
    theme_light()
```


The total moose abundance coverage was lower than nominal for all models:

```{r}
#| echo: false
#| output: asis
dd |>
    group_by(true_model, model) |>
    summarize(pi_90 = mean(pi_90), pi_95 = mean(pi_95), .groups = "keep") |>
    knitr::kable()
```

The bias (estimate - true) in total abundance:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = bias, col = model)) +
    geom_boxplot() +
    theme_light()
```

Relative bias (standardized by true abundance):

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = rel_bias, col = model)) +
    geom_boxplot() +
    theme_light()
```

Cell level coverage was higher than nominal:

```{r}
#| echo: false
dd |>
    ggplot(aes(x = true_model, y = coverage_95, col = model)) +
    geom_boxplot() +
    theme_light()
dd |>
    ggplot(aes(x = true_model, y = coverage_90, col = model)) +
    geom_boxplot() +
    theme_light()
```

