---
title: "Simulation Approach"
subtitle: "Moose Abundance Estimation"
date: "`r Sys.Date()`"
author: "Péter Sólymos"
toc: true
format:
  html: 
    html-math-method: katex
    self-contained: true
  pdf: default
---

```{r}
#| include: false
library(knitr)
if (interactive()) {
    devtools::load_all(".")
} else {
    devtools::load_all("../..")
}
set.seed(1)
get_stats <- function(PI, true_vals) {
    is_srv <- PI$data$srv
    cell_q <- t(apply(
        PI$boot_full,
        1,
        quantile,
        c(0.025, 0.05, 0.95, 0.975)
    ))
    in_95 <- true_vals %[]% cell_q[, c(1, 4)]
    in_90 <- true_vals %[]% cell_q[, c(2, 3)]
    tot <- colSums(PI$boot_full)
    tot_mean <- mean(tot)
    tot_q <- quantile(tot, c(0.025, 0.05, 0.95, 0.975))
    c(
        true_total = sum(true_vals),
        mean = tot_mean,
        tot_q,
        coverage_90 = sum(in_90[!is_srv]) / sum(!is_srv),
        coverage_95 = sum(in_95[!is_srv]) / sum(!is_srv)
    )
}

```

# Preamble

```{r}
suppressPackageStartupMessages({
    library(dplyr)
    library(ggplot2)
    library(intrval)
})
```

# Simulating data from count distributions

There are some functions in the package that allow us to simulate data.

`sim_rsf()` generates a spatial random field represented by (x,y,z) pairs of
values: coordinates (x, y), and the value of the RSF functions that is spatially
correlated to imitate real world data sets. The grid has a default size
of 50 x 50 = 2500 cells.

```{r}
m <- sim_rsf()
summary(m)

m |>
    ggplot(aes(x = x, y = y, z = z)) +
    geom_contour_filled() +
    theme_light()
```

We can stratify the grid using `sim_strat()`.
`threshold` is the cutoff value for the low (0 values, < threshold)
and high (1 values, >= threshold); `p_misclass` is the misclassification 
probability (0 = perfect, 1 = random) indicating what proportion of the 
classes are misclassified (taking the opposite value).

NOTE: we can probably make this more realistic if the misclassification only
happens in an interval around the threshold and not over the full range of the RSF.

```{r}
m <- m |>
    sim_strat(threshold = 0, p_misclass = 0.1)
summary(m)
table(m$stratum, m$z > 0)
1 - sum(diag(table(m$stratum, m$z > 0))) / nrow(m)

m |>
    ggplot(aes(x = x, y = y, z = z)) +
    geom_contour_filled() +
    geom_point(aes(x = x, y = y, col = as.factor(stratum)), m) +
    theme_light()
```

The `sim_expected()` function will add the column `mu` for the expected Moose
abundance calculated as `mu = exp(intercept + z * slope)` based on the RSF value.
The other column, `phi.zi` is the probability of nonzero counts for ZI and
Huardle models later (phi1 = plogis(intercept1 + slope1 * sim_rsf$z))

```{r}
m <- m |>
    sim_expected(
        intercept = -1, slope = 0.5,
        intercept1 = -1, slope1 = 0.25
    )
summary(m)

m |>
    ggplot(aes(x = x, y = y, z = mu)) +
    geom_contour_filled() +
    theme_light()
```

Finally, we simulate the counts according to one of the chosen count distribution
using `sim_counts()`.
The `theta.nb` sets the dispersion parameter for the Negative Binomial
(Poisson when left as `NULL`). The `type` argument takes a value
`"none"`, `"zeroinfl"` or `"hurdle"`.

|  Distribution    |  `theta.nb`  |  `type`      |
|------------------|--------------|--------------|
| Poisson          |  `NULL`      |     `"none"`      |
| NegBin           |  Not `NULL`  |    `"none"`       |
| ZI Poisson       |  `NULL`      |    `"zeroinfl"`       |
| ZI NegBin        |  Not `NULL`  |     `"zeroinfl"`      |
| Hurdle Poisson   |  `NULL`      |    `"hurdle"`       |
| Hurdle NegBin    |  Not `NULL`  |    `"hurdle"`       |

The `sim_counts()` function add the `MOOSE_TOTA` to our data frame:


```{r}
m <- m |>
    sim_counts(theta.nb = 0.1, type = "zeroinfl")
summary(m)
table(m$MOOSE_TOTA)
plot(table(m$MOOSE_TOTA))

m |>
    ggplot(aes(x = x, y = y, z = mu)) +
    geom_contour_filled() +
    geom_point(aes(x = x, y = y, col = MOOSE_TOTA), m[m$MOOSE_TOTA > 0, ]) +
    theme_light()
```

# Analyze simulated data

```{r}
mc_options(B = 100)
switch_response("total")

# sample 100 cells
surv_id <- sample(nrow(m), 100)
m$srv <- FALSE
m$srv[surv_id] <- TRUE
m$AREA_KM <- 16

# fit models
a <- c(
    "P" = "P",
    "NB" = "NB",
    "ZIP" = "ZIP",
    "ZINB" = "ZINB"
)
ML <- lapply(a, function(i) {
    mc_fit_total(m, vars = "z", zi_vars = "z", dist = i)
})

# get PIs
PIL <- list()
for (i in a) {
    PIL[[i]] <- mc_predict_total(
        model_id = i,
        ml = ML,
        x = m,
        do_boot = TRUE,
        do_avg = FALSE
    )
}

# true abundance
sum(m$MOOSE_TOTA)

# Count model estimates
sapply(PIL, get_stats, true_vals = m$MOOSE_TOTA)

# Gassaway
mc_gassaway(
    y1 = m$MOOSE_TOTA[m$srv & m$stratum == 0],
    y2 = m$MOOSE_TOTA[m$srv & m$stratum == 1],
    N1 = sum(1 - m$stratum),
    N2 = sum(m$stratum)
)
```
